---
title: "2-dplyr"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(nycflights13))
suppressPackageStartupMessages(library(tibble))
```

## Dplyr: Data manipulation

There are six main verbs for data manipulations:

* Filter: to pick some rows based on a boolean test.
* Select: to pick some columns.
* Arrange: to arrange the rows in a specified order.
* Mutate: to create new variables.
* Summarise: to collapse a data frame in one value.
* Group_by: to group a dataframe by some unique values of the variables. 

Let's start with the tibble: a modern take on the dataframe.

```{r}
flights
```

Any duplicated rows? 

```{r}
dplyr::distinct(flights)
```

A tibble won't change strings to factor and has a handy print output. To change to tibble, use
__as_tibble()__ function. 

# Filter

Let's start with the first verb: filter. Subset the rows with a logical condition. Let's take all the flights in 2013 in the first of January. Note: the conditions are taken as intersection. 

```{r}
flights %>%
  filter(month == 1, day == 1)
```

To take union of logical conditions, put the union in one logical condition. Let's say all the flights in June, July and August

```{r}
flights %>%
  filter(month %in% c(6,7,8))
```

# Arrange

The verb to arrange the rows in a particular order. The default is in ascending order. If more variables
are added, they'll be used hierarchically. With __desc()__ , it's organized in descending order.


```{r}
flights %>%
  arrange(desc(arr_delay), desc(dep_delay))
```

## Exercise

Sort flights to find the most delayed flights. Find the flights that left earliest. Most delayed:

```{r}
flights %>%
  arrange(desc(arr_delay), desc(dep_delay))
```

Left earlies:

```{r}
flights %>%
  arrange(dep_delay)
```

# Select

Select: the verb to select some variables of the data frame

```{r}
flights %>%
  select(ends_with("delay"), starts_with("arr"), starts_with("sched"), starts_with("dep"))
```

# Mutate

__Mutate()__: the verb to create new variables with functions of existing variables. 

```{r}
flights %>%
  mutate(gain = arr_delay - dep_delay,
       speed = distance / air_time * 60)
```

If you only want to keep the new variables, use __transmute()__ : 

```{r}
transmute(flights,
          dep_time, # equivalent to keep in new dataframe that transmutes creates
          # dep_time is hour and then minutes. Last two digits, minutes
          hour = dep_time %/% 100, # keep the integer part of the division
          minute = dep_time %% 100 # keep the remainder part of division
)
```

# Rename

__rename()__ : the verb to rename variables:

```{r}
flights %>%
  rename(departure_time = dep_time)
```

# Summarise

__sumarise()__ : the verb to collapse a data frame into one dimension. 

```{r, include=FALSE}
flights %>%
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE))
```

# Group_by

__group_by()__ : the verb to group variables according to some unique combination of values, and then perform operations on these groups.

```{r}
flights %>%
  filter(!is.na(arr_delay), !is.na(dep_delay)) %>%
  group_by(year, month, day, origin) %>% # For each day of the year, for each one of the 3 airports, compute the mean delay in both arrival 
  # and departure
  summarise_each(funs(mean), avg_arr_delay = arr_delay, avg_dep_delay = dep_delay) %>%
  mutate(total_avg_delay = avg_arr_delay + avg_dep_delay)
```



