---
title: "Strings in R"
output: html_notebook
---

```{r}
suppressPackageStartupMessages(library(stringr))
library(stringi)
library(readr)
```

# String Basics

```{r}
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
writeLines(string2)
```

To get the length of a string:

```{r}
str_length(string2)
```

To combine strings: 

```{r}
str_c("combine","strings", sep = "-")
str_c("prefix-", c("a", "b", "c"), "-suffix") # Note the recycling
```
```{r}
name <- "Hadley"
time_of_day <- "morning"
birthday <- FALSE

str_c("Good ", time_of_day, " ", name,
  if (birthday) " and HAPPY BIRTHDAY",
  "."
)

```

```{r}
x <- c("Apple", "Banana", "Pear")
str_sub(x, start = 1, end = 3)
# negative numbers count backwards from end
str_sub(x, start = -3, end = -1)
# Modify string
(str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1)))
```
```{r}
str_sort(x)
```
```{r}
y <- "My name is"
z <- "Genghis"
paste0(y," ",z)
str_c(y, z, sep = " ")
```

Use __str_length()__ and __str_sub()__ to extract the middle character from a character vector.

```{r}
w <- c("Genghis","Khan","Marco Polo")
mid <- str_length(w)/2
w %>%
  str_sub(start = mid, end = mid)
```

# Matching patterns with Regular Expressions

To see how a regular expression plays out with some character vector, use __str_view()__

```{r}
x <- c("apple", "banana", "pear")
str_view_all(x, "an")
```

The dot is used to represent any possible character. So .a. will match any three letter sequence with an 'a' in between. 

```{r}
str_view_all(x, ".a.")
```

The caret ^ will make sure that you're matching the start of the string. Such that, ^a will match if the first letter is a

```{r}

```

```{r}
x <- c("apple", "banana", "pear")
str_view(x, "^p")
```

Match at the end, use $.

```{r}
str_view(x, "a$")
```

To force a regular expression to only match a complete string, anchor it with both ^ and $. Remember, each component of the string vector is treated as a different string. 

```{r}
x <- c("apple pie", "apple", "apple cake")
str_view(x, "^apple$")
```

# Putting it together

To determine if a character vector matches a pattern, use str_detect(). It returns a logical vector the same length as the input:

```{r}
x <- c("apple", "banana", "pear")
str_detect(x, "e")
```

```{r}
words <- c("avocado","devil","advocate","lawyerly","Coase","Economics")
str_detect(words, "(^C)|(c)" ) # Starts with C or has a c in it. 
```

```{r}
str_detect(words, "^([aeiou])|([aeiou])$") #starts or ends with a lowercase vocal
```

To select elements that have a common pattern: 

```{r}
str_subset(words, "[^aeiou]$") # Words that do not end up in a vocal
```

```{r}
str_subset(words, "[sv]") #words that contain either an s or a v, or both
```

To count the number of matches per string:

```{r}
str_count(words, "[sv]")
```

Find all words that start or end with a consonant. 

```{r}
str_subset(words, "[^aeiou]$")
```

Find all words that start with a vowel and end with a consonant.

```{r}
str_subset(words, "(^[aeiou]) & ([^aeiou])")
```

Are there any words that contain at least one of each different vowel?

```{r}
str_subset(words, "(a+) & (e+) & (i+) & (o+) & (u+)")
```

What word has the highest number of vowels? What word has the highest proportion of vowels? 

```{r}
count <- str_count(words, "[aeiou]")
len <- str_length(words)
words[which.max(count/len)]
```

# Extract matches

with str_extract, we can extrac what is the match.

```{r}
setwd(dir = "/Users/david_salazarv/Desktop/Tidyverse")
sentences <- read_lines("harvard-sentences.txt")
head(sentences)
```


```{r}
colours <- c("red", "orange", "yellow", "green", "blue", "purple")
colour_match <- str_c(colours, collapse = "|")
colour_match
```

Now we can select the sentences that contain a colour, and then extract the colour to figure out which one it is:

```{r}
has_colour <- str_subset(sentences, colour_match)
matches <- str_extract(has_colour, colour_match)
head(matches)
```

Extract all the matches for every single string:

```{r}
head(str_extract_all(has_colour, colour_match, simplify = TRUE))
```

str_replace() and str_replace_all() allow you to replace matches with new strings:

```{r}
x <- c("apple", "pear", "banana")
str_replace_all(x, "[aeiou]", "-")
```

```{r}
x <- c("1 house", "2 cars", "3 people")
str_replace_all(x, c("1" = "one", "2" = "two", "3" = "three"))
```

