---
title: "R Notebook"
output: html_notebook
---

# Dates and Times

```{r, warning=FALSE}
library(lubridate)
library(nycflights13)
library(dplyr)
library(ggplot2)
```

Let's work with the flights dataset:

```{r}
flights %>% 
  select(year, month, day, hour, minute)
```

As you can see, month, day, and hour are in separate columns; let's put them together in a column with the appropriate type. To combine separate numbers into datetimes, use make_datetime()

```{r}
datetimes <- flights %>% 
  mutate(departure = make_datetime(year = year, month = month, day = day, 
                                   hour = hour, min = minute))
datetimes %>%
  select(year, month, day, hour, minute, departure)
```

```{r}
flights %>%
  select(arr_time)
```


With a little work, we can also create arrival times for each flight in flights. I’ll then clean up the data a little.

```{r}
(datetimes <- datetimes %>% 
  mutate(arrival = make_datetime(
    year = year, 
    month = month, 
    day = day, 
    hour = arr_time %/% 100,  # integer part of division by 100
    min =  arr_time %% 100  # decimal part of division by hundred
  )) %>% 
  filter(!is.na(departure), !is.na(arrival)) %>% 
  select(
    departure, arrival, dep_delay, arr_delay, carrier, tailnum, 
    flight, origin, dest, air_time, distance
  )
)
```

To parse character strings as dates, identify the order in which the year, month, and day appears in your dates. Now arrange “y”, “m”, and “d” in the same order. This is the name of the function in lubridate that will parse your dates. For example,

```{r}
ymd("20170131")
mdy("January 31st, 2017")
dmy("31-1-2017")
```

R now recognizes that our departure and arrival variables contain datetime information, and it saves the variables in the POSIXct format, a common way of representing dates and times. The POSIXct format has many advantages. You can display the same date time in any time zone by changing its tzone attribute (more on that later), and R can recognize when two times displayed in two different time zones refer to the same moment.

Best of all, you can change a datetime by adding or subtracting seconds from it.

This gives us a way to calculate the scheduled departure and arrival times of each flight in flights.

```{r}
datetimes %>% 
  mutate(scheduled_departure = departure - dep_delay * 60,
         scheduled_arrival = arrival - arr_delay * 60) %>% 
  select(scheduled_departure, dep_delay, departure, 
         scheduled_arrival, arr_delay, arrival)
```

## Difftimes

A difftime class object records a span of time in one of seconds, minutes, hours, days, or weeks. R creates a difftime whenever you subtract two dates or two datetimes.

```{r}
(day1 <- ymd("2000-01-01") - ymd("1999-12-31"))
```

Durations behave like difftimes, but are a little more user friendly. To make a duration, choose a unit of time, make it plural, and then place a “d” in front of it. 

```{r}
dseconds(1)
dhours(1)
dweeks(1)
dyears(1)
```

A better, and clearer way to do arithmetic with dates and times:

```{r}
(datetimes <- datetimes %>% 
  mutate(scheduled_departure = departure - dminutes(dep_delay),
         scheduled_arrival = arrival - dminutes(arr_delay)) %>% 
  select(scheduled_departure, dep_delay, departure, 
         scheduled_arrival, arr_delay, arrival, 
         carrier, tailnum, flight, origin, dest, air_time, distance))
```

```{r}
datetimes %>% 
  ggplot(aes(scheduled_departure)) + 
  geom_histogram(binwidth = 86400) # 86400 seconds = 1 day
```

You can extract the year, month, day of the year (yday), day of the month (mday), day of the week (wday), hour, minute, second, and time zone (tz) of any date or datetime with lubridate’s accessor functions. Use the function that has the name of the unit you wish to extract.

```{r}
(datetime <- ymd_hms("2007-08-09 12:34:56", tz = "America/Los_Angeles"))
#> [1] "2007-08-09 12:34:56 PDT"
year(datetime)
#> [1] 2007
month(datetime)
#> [1] 8
yday(datetime)
#> [1] 221
mday(datetime)
#> [1] 9
wday(datetime)
#> [1] 5
hour(datetime)
#> [1] 12
minute(datetime)
#> [1] 34
second(datetime)
#> [1] 56
tz(datetime)
#> [1] "America/Los_Angeles"
```

For both month() and wday() you can set label = TRUE to return the name of the month or day of the week. Set abbr = TRUE to return an abbreviated version of the name, which can be helpful in plots.


```{r}
datetimes %>% 
  transmute(weekday = wday(scheduled_departure, label = TRUE)) %>% 
  filter(!is.na(weekday)) %>% 
  ggplot(aes(x = weekday)) +
    geom_bar()
```

```{r}
datetimes %>% 
  transmute(hour = hour(scheduled_departure)) %>% 
  filter(!is.na(hour)) %>% 
  ggplot(aes(x = hour)) +
    geom_bar()
```

When should you depart if you want to minimize your chance of delay? The results are striking. On average, flights that left on a Saturday arrived ahead of schedule.

```{r, warning=FALSE}
datetimes %>% 
  mutate(weekday = wday(scheduled_departure, label = TRUE)) %>% 
  filter(!is.na(weekday)) %>% 
  group_by(weekday) %>% 
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = weekday, y = avg_delay)) +
    geom_bar(stat = "identity") +
    theme_minimal()
```

```{r, warning=FALSE}
datetimes %>% 
  mutate(hour = hour(scheduled_departure)) %>% 
  filter(!is.na(hour)) %>% 
  group_by(hour) %>% 
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = hour, y = avg_delay)) +
    geom_bar(stat = "identity")
```

Now, let's try to get the average delay per month of the year

```{r, warning=FALSE}
datetimes %>% 
  mutate(month = month(scheduled_departure, label = TRUE)) %>% 
  filter(!is.na(month)) %>% 
  group_by(month) %>% 
  summarise(avg_delay = mean(arr_delay, na.rm = TRUE)) %>% 
  ggplot(aes(x = month, y = avg_delay)) +
    geom_bar(stat = "identity") +
    theme_minimal()
```

